---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---
Use the Peruvian data, and prepare a code in R to answer:
- Is the whole network connected? If disconnected, keep the giant component to answer all the remaining questions. If connected, use the whole graph for the remaining questions.
- Who are the nodes that represent the vertices in the diameter?
- Make a plot that highlights the nodes with the highest betweenness.
- Make a DataFrame with three the centrality measures, and plot those three values in a scatterplot.

```{r}
rm(list = ls())
GitLink='https://github.com/DACSS690C-CSSMETHODS/HW2/raw/refs/heads/main/graph_FigueData.graphml'
```

```{r}
library(igraph)
peru=read_graph(GitLink,format='graphml')
```

```{r}
summary(peru)
```
This is a summary of my directed (D) weighted (W) graph, with 37 nodes and 266 edges.

Adding names to my nodes:
```{r}
V(peru)$name=V(peru)$id
```


Question 1: Is the whole network connected? If disconnected, keep the giant component to answer all the remaining questions. If connected, use the whole graph for the remaining questions.
```{r}
is_connected(peru)
```
It looks like the network is disconnected. 

Question 2: Who are the nodes that represent the vertices in the diameter?

Option A method:
```{r}
isolates = which(degree(peru,mode='in') + degree(peru,mode='out')==0 )
peru_noIsolates = delete_vertices(peru, isolates)
## result
summary(peru_noIsolates)
```

Now I can find nodes that represent the vertices in the diameter
```{r}
diameter(peru_noIsolates)
```
The diameter of a graph is the length of the longest shortest path between any two vertices. The output shows that the diameter is 4.

```{r}
farthest_vertices(peru_noIsolates)
```
Finding the nodes whose geodesic represent the diameter: Michell and Wong Lu


```{r}
get_diameter(peru_noIsolates)
```
The nodes that represent the vertices in the diameter of the graph peru_noIsolates are:
Vertices: Michell      Miro Quesada Romero       Bentin       Wong Lu   
These vertices form the longest shortest path in the graph, which has a length of 4.

Question 3: Make a plot that highlights the nodes with the highest betweenness.
```{r}
sort(betweenness(peru_noIsolates,normalized=T))
```

```{r}
#getting betweenness centrality for all nodes and sort it
betweenness_scores = betweenness(peru_noIsolates, directed = TRUE, weights = E(peru_noIsolates)$weight, normalized = TRUE)
sorted_betweenness_scores = sort(betweenness_scores, decreasing = TRUE)

#defining a threshold for top betweenness nodes
top_n = 5
top_betweenness_threshold = sorted_betweenness_scores[top_n]

#indentifying the nodes with the highest betweenness centrality
top_betweenness_nodes = V(peru_noIsolates)[betweenness_scores >= top_betweenness_threshold]

#colors
vcol = rep("grey80", vcount(peru_noIsolates))
vcol[top_betweenness_nodes] = "red"

#plotting
plot(peru_noIsolates, vertex.color = vcol, layout = layout.fruchterman.reingold, 
     vertex.label.cex = 0.8, vertex.size = 5, edge.arrow.size = 0.5)
```

Question 4: Make a DataFrame with three the centrality measures, and plot those three values in a scatterplot.
```{r}
#three the centrality measures
rounding=2
degr=round(degree(peru_noIsolates,normalized=T),rounding)
close=round(closeness(peru_noIsolates,normalized=T),rounding)
betw=round(betweenness(peru_noIsolates,normalized=T),rounding)
#making the dataframe
DFCentrality=as.data.frame(cbind(degr,close,betw),stringsAsFactors = F)
names(DFCentrality)=c('Degree','Closeness','Betweenness')
```

```{r}
DFCentrality$person=row.names(DFCentrality)
row.names(DFCentrality)=NULL
```

```{r}
#plotting scatterplot
library(ggplot2)
library(ggrepel)
ggplot(DFCentrality, aes(x=Betweenness, y=Closeness)) + 
    theme_classic() + geom_point() +
    geom_label_repel(aes(label=person,fill=Degree),color='white') +
    scale_fill_gradient2(low = "red", mid='grey50',high ="darkgreen",
                           midpoint = 0.3)
```

Use the Seattle data, and prepare a code in R to answer:
- Is the whole network connected? If disconnected, keep the giant component to answer all the remaining questions. If connected, use the whole graph for the remaining questions.
- How many nodes should be deleted to split this network?
- Make a DataFrame with three the Link Analysis measures Hubs, Authorities, and PageRank, and plot those three values in a scatterplot.
```{r}
rm(list = ls())
GitLink='https://github.com/DACSS690C-CSSMETHODS/HW2/raw/refs/heads/main/graph_edgelist.graphml'
```

```{r}
library(igraph)
seattle=read_graph(GitLink,format='graphml')
```

```{r}
summary(seattle)
```
This is a summary of my directed (D) weighted (W) graph, with 46 nodes and 588 edges.

Adding names to my nodes:
```{r}
V(seattle)$name=V(seattle)$id
```

Question 1: Is the whole network connected? If disconnected, keep the giant component to answer all the remaining questions. If connected, use the whole graph for the remaining questions.
```{r}
is_connected(seattle)
```
It looks like the network is connected.

Question 2: How many nodes should be deleted to split this network?
```{r}
vertex_connectivity(seattle)
```
0 nodes should be deleted.

Question 3: Make a DataFrame with three the Link Analysis measures Hubs, Authorities, and PageRank, and plot those three values in a scatterplot.
```{r}
#my three link analysis measures
Lk_hub=hub_score(seattle)$vector
Lk_auth=authority_score(seattle)$vector
Lk_page=page_rank(seattle)$vector
#making the dataframe
seattle_LINK=as.data.frame(cbind(Lk_hub,Lk_auth,Lk_page))
seattle_LINK
```

```{r}
#scatterplot time
seattle_LINK$person=row.names(seattle_LINK)

ggplot(seattle_LINK, aes(x=Lk_hub, y=Lk_auth)) + 
    theme_classic() + geom_point() +
    geom_label_repel(aes(label=person,fill=Lk_page),color='white') +
    scale_fill_gradient2(low = "red", mid='grey50',high ="darkgreen",
                           midpoint = 0.05)
```

Use the Fifa Data, projecting only the countries, and report:
- Is the whole network connected? If disconnected, keep the giant component to answer all the remaining questions. If connected, use the whole graph for the remaining questions.
- What is the density of this network?
- Make a DataFrame with three the centrality measures, and plot those three values in a scatterplot.
```{r}
rm(list = ls())
GitLink='https://github.com/DACSS690C-CSSMETHODS/HW2/raw/refs/heads/main/country_projection.graphml'
```

```{r}
library(igraph)
fifa=read_graph(GitLink,format='graphml')
```

```{r}
summary(fifa)
```
This is a summary of my undirected graph, with 32 nodes and 305 edges.

Adding names to my nodes:
```{r}
V(fifa)$name=V(fifa)$id
```

Question 1: Is the whole network connected? If disconnected, keep the giant component to answer all the remaining questions. If connected, use the whole graph for the remaining questions.
```{r}
is_connected(fifa)
```
It looks like the network is connected.

Question 2: What is the density of this network?
```{r}
edge_density(fifa)
```

Question 3: Make a DataFrame with three the centrality measures, and plot those three values in a scatterplot.
```{r}
#my three centrality measures
rounding=2
degr=round(degree(fifa,normalized=T),rounding)
close=round(closeness(fifa,normalized=T),rounding)
betw=round(betweenness(fifa,normalized=T),rounding)
#making the dataframe
DFCentrality=as.data.frame(cbind(degr,close,betw),stringsAsFactors = F)
names(DFCentrality)=c('Degree','Closeness','Betweenness')
```

```{r}
DFCentrality$person=row.names(DFCentrality)
row.names(DFCentrality)=NULL
```

```{r}
#plotting
library(ggplot2)
library(ggrepel)
ggplot(DFCentrality, aes(x=Betweenness, y=Closeness)) + 
    theme_classic() + geom_point() +
    geom_label_repel(aes(label=person,fill=Degree),color='white') +
    scale_fill_gradient2(low = "red", mid='grey50',high ="darkgreen",
                           midpoint = 0.3)
```


